Pattern matching for (231,213) avoiding permutations

### 2015-01-30 article.tex
correction : remove the special case of last element for stripe. correction : proof for pattern matching for two permutation avoiding the motif to adapt with the definition of stripe. correction : definition of a segment in pattern mathing with motif avoiding hte motif to adapt with the definition of stripe.

		A \textbf{Stripe} of a permutation
		 is a mapping into \textbf{up step} (noted $\ustep$) 
		or a \textbf{down step} (noted $\dstep$).
		An element $i$ of a permutation $p$ is an up step if and only if $i$ is not the last element and $p(i)<p(i+1)$.
		An element $i$ of a permutation $p$ is a down step if and only if $i$ is not the last element and $p(i)>p(i+1)$.	
		By convention the last element is mapped as the second to last element.\\ 	
		Formally a Stripe is a function $\mathcal{S} : [n] \to \{\dstep,\ustep\}$ define as :\\
		$ \mathcal{S}(p,i) = \left\{ 
		  \begin{array}{l l}
		    \ustep & \quad \text{if $1 \leq i<n$ and $p(i+1)-p(i)>0$ }\\
		    \dstep & \quad \text{if $1 \leq i<n$ and $p(i+1)-p(i)<0$ }\\
		     \mathcal{S}(p,n-1) & \quad \text{if $i=n$}\\
		  \end{array} \right. $\\

		By extension we define  $\mathcal{S} : \Perm{n} \to \{\dstep,\ustep\}^{n}$ such as :
		$\mathcal{S}(p)=\mathcal{S}(p,1)\mathcal{S}(p,2)...\mathcal{S}(p,n)$.
		We call $\mathcal{S}(p)$ the stripe of $p$.
		For example the stripe of the permutation $(12,11,10,1,2,3,4,9,8,5,6,7)$, 
		is $\dstep,\dstep,\dstep,\ustep,\ustep,\ustep,\ustep,\dstep,\dstep,\ustep,\ustep,\ustep$.\\

@@

		Formally we say that the stripe of $\pmotif \in \Perm{m}$ is a motif
		of a stripe of $\ptext \in \Perm{n}$ if and only if there exist an increasing function $\phi : [|\pmotif|] \rightarrow [|\pmotif|]$
		such that : \\
		$\forall i \in [|\pmotif|]$, $\mathcal{S}(\pmotif,i) = \mathcal{S}(\ptext,\phi(i))$\\
		If a stripe $R_2$ is a motif of a stripe $R_1$,
		it means that there exists
		a subsequence of $R_1$ such as
		this sequence is equal to $R_2$.
%		We say that a set is a match 
%		if and only if there exists a function such as define above
%		and the set is the set of image of this function.\\
		For example, the stripe $(\dstep_1,\dstep_2,\dstep_3,\dstep_4,\ustep_1,\ustep_2)$ is a motif of  the stripe \\
		$(\dstep_1,\dstep_2,\dstep_3,\ustep_1,\ustep_2,\ustep_3,\ustep_4,\dstep_4,\dstep_5,\ustep_5,\ustep_6,\ustep_7)$ ,
		because we can match \\
		$(\dstep_1,\dstep_2,\dstep_3,\dstep_4,(\ustep_1,\ustep_2)$ to $(\dstep_1,\dstep_2,\dstep_3,\dstep_4,\ustep_5,\ustep_6)$.
		\begin{figure}[h!]
			\begin{displaymath}
				\xymatrix@=0.3cm{
					\dstep_1 \ar[dd] & \dstep_2 \ar[dd] & \dstep_3 \ar[dd] & \dstep_4\ar[ddrrrr]  & \ustep_1 \ar[ddrrrrr]& \ustep_2 \ar[ddrrrrr]  \\
					\\
					\dstep_1 & \dstep_2 & \dstep_3 & \ustep_1 & \ustep_2 & \ustep_3 & \ustep_4 & \dstep_4 & \dstep_5 & \ustep_5 & \ustep_6 & \ustep_7 & \\
				}
			\end{displaymath}     
		\end{figure}		
		But the stripe $(\dstep,\ustep,\dstep,\dstep,\dstep)$ is not a motif of the stripe\\ $(\dstep_1,\dstep_2,\dstep_3,\ustep_1,\ustep_2,\ustep_3,\ustep_4,\dstep_4,\dstep_5,\ustep_5,\ustep_6,\ustep_7)$.\\

@@

		\textbf{Corollary \numl{stripe} \thenum}  If $p \in \Av{n}{213,231}$ its stripe can be define such as, \\
		$\mathcal{S}(p,i) = \left\{ 
  		\begin{array}{l l}
   			\ustep & \quad \text{if $p(i)$ is the minimal element of $p[i:]$}\\    
   			\dstep & \quad \text{if $p(i)$ is the maximal element of $p[i:]$}\\
  		 	\mathcal{S}(p,n-1) & \quad \text{if $i=n$}\\
	 	\end{array} \right. 
	 	$\\

@@

		\textbf{Proof} \\
		$\Rightarrow$ Suppose that $\pmotif$ is a motif of $\ptext$
		by subsequence $s$.
		By definition $s$ normalized is equal to
		$\pmotif$ so they have the same stripe.\\ 
		$\Leftarrow$ 
		Let $\ptext \in \Perm{n}$, 
		$R_1$ be its stripe, 
		$\pmotif \in \Perm{m}$
		and $R_2$ its stripe.
		Let $\phi$ be the function
		that match $R_2$ to $R_1$.
		We define $\varphi$ as, 
		$\forall x \in [m]$,
		$\varphi(x) = \phi(x)$.
		Suppose that $\exists x,y \in [|\pmotif|]$, 
		$\pmotif(x)<\pmotif(y)$ 
		and $\ptext(\varphi(x))>\ptext(\varphi(y))$,
		without lost of generality we can suppose that $x<y$,
		Let see all the case possible that can make this 
		hypothesis possible :
		\begin{itemize}

			\item if $\pmotif(x)<\pmotif(y)$, 
			$\mathcal{S}(\pmotif,x)=\ustep$ 
			and $\mathcal{S}(\pmotif,y)=\ustep$
			(if x and y are increasing in $\pmotif$),
			then $\mathcal{S}(\ptext,\varphi(x))=\ustep$ and $\mathcal{S}(\ptext,\varphi(y))=\ustep$,
			(the match are also increasing).
			As $x<y$ and as $\varphi$ is an increasing function, 
			$\varphi(x) < \varphi(y)$.
			And thus, by \textbf{Lemma \ref{upup}},
			$\ptext(\varphi(x)) < \ptext(\varphi(y))$ .
			Which contradict our hypothesis,
			so this case is not possible.

@@

@@

			For the algorithm we introduce a decomposition,
			we decompose the motif into
			adjacent element of same step in the stripe 
			(called such sequence segment),
			and label them form right to left
			with increasing number.
			Let's call this decomposition $S$.
			For each segment, we call important 
			the leftmost element denoted by $IMP$.
			For example, given the motif $(12,11,10,1,2,3,4,9,8,5,6,7)$,
			$S(1)=(5,6,7)$, $IMP(1)=5$, 
			$S(2)=(9,8)$, $IMP(2)=9$, 
			$S(3)=(1,2,3,4)$, $IMP(3)=1$, 
			$S(4)=(12,11,10)$ and $IMP(4)=12$.\\

>>>>

		A \textbf{Stripe} of a permutation
		is a mapping into \textbf{up step} (noted $\ustep$) 
		or a \textbf{down step} (noted $\dstep$).
		An element $i$ of a permutation $p$ is an up step if and only if $i$ is not the last element and $p(i)<p(i+1)$.
		An element $i$ of a permutation $p$ is a down step if and only if $i$ is not the last element and $p(i)>p(i+1)$.\\	
		Formally a Stripe is a function $\mathcal{S} : [n-1] \to \{\dstep,\ustep\}$ define as :\\
		$ \mathcal{S}(p,i) = \left\{ 
		  \begin{array}{l l}
		    \ustep & \quad \text{if $1 \leq i<n$ and $p(i+1)-p(i)>0$ }\\
		    \dstep & \quad \text{if $1 \leq i<n$ and $p(i+1)-p(i)<0$ }\\
		  \end{array} \right. $\\

		By extension we define  $\mathcal{S} : \Perm{n} \to \{\dstep,\ustep\}^{n}$ such as :
		$\mathcal{S}(p)=\mathcal{S}(p,1)\mathcal{S}(p,2)...\mathcal{S}(p,n-1)$.
		We call $\mathcal{S}(p)$ the stripe of $p$.
		For example the stripe of the permutation $(12,11,10,1,2,3,4,9,8,5,6,7)$, 
		is $\dstep,\dstep,\dstep,\ustep,\ustep,\ustep,\ustep,\dstep,\dstep,\ustep,\ustep$.\\

@@

		Formally we say that the stripe of $\pmotif \in \Perm{m}$ is a motif
		of a stripe of $\ptext \in \Perm{n}$ if and only if there exist an increasing function $\phi : [|\pmotif|] \rightarrow [|\pmotif|]$
		such that : \\
		$\forall i \in [|\pmotif|]$, $\mathcal{S}(\pmotif,i) = \mathcal{S}(\ptext,\phi(i))$\\
		If a stripe $R_2$ is a motif of a stripe $R_1$,
		it means that there exists
		a subsequence of $R_1$ such as
		this sequence is equal to $R_2$.
%		We say that a set is a match 
%		if and only if there exists a function such as define above
%		and the set is the set of image of this function.\\
		For example, the stripe $(\dstep_1,\dstep_2,\dstep_3,\dstep_4,\ustep_1)$ is a motif of  the stripe \\
		$(\dstep_1,\dstep_2,\dstep_3,\ustep_1,\ustep_2,\ustep_3,\ustep_4,\dstep_4,\dstep_5,\ustep_5,\ustep_6)$ ,
		because we can match \\
		$(\dstep_1,\dstep_2,\dstep_3,\dstep_4,\ustep_1)$ to $(\dstep_1,\dstep_2,\dstep_3,\dstep_4,\ustep_5,\ustep_6)$.
		\begin{figure}[h!]
			\begin{displaymath}
				\xymatrix@=0.3cm{
					\dstep_1 \ar[dd] & \dstep_2 \ar[dd] & \dstep_3 \ar[dd] & \dstep_4\ar[ddrrrr]  & \ustep_1 \ar[ddrrrrr]\\
					\\
					\dstep_1 & \dstep_2 & \dstep_3 & \ustep_1 & \ustep_2 & \ustep_3 & \ustep_4 & \dstep_4 & \dstep_5 & \ustep_5 & \ustep_6 \\
				}
			\end{displaymath}     
		\end{figure}		
		But the stripe $(\dstep,\ustep,\dstep,\dstep,\dstep)$ is not a motif of the stripe\\ $(\dstep_1,\dstep_2,\dstep_3,\ustep_1,\ustep_2,\ustep_3,\ustep_4,\dstep_4,\dstep_5,\ustep_5,\ustep_6,\ustep_7)$.\\

@@

		\textbf{Corollary \numl{stripe} \thenum}  If $p \in \Av{n}{213,231}$ its stripe can be define such as, \\
		$\mathcal{S}(p,i) = \left\{ 
  		\begin{array}{l l}
   			\ustep & \quad \text{if $p(i)$ is the minimal element of $p[i:]$}\\    
   			\dstep & \quad \text{if $p(i)$ is the maximal element of $p[i:]$}\\
	 	\end{array} \right. 
	 	$\\

@@

		\textbf{Proof} \\
		$\Rightarrow$ Suppose that $\pmotif$ is a motif of $\ptext$
		by subsequence $s$.
		By definition $s$ normalized is equal to
		$\pmotif$ so they have the same stripe.\\ 
		$\Leftarrow$ 
		Let $\ptext$ be a permutation avoiding $(213,231)$, 
		$R_1$ be its stripe, 
		$\pmotif$  be a permutation avoiding $(213,231)$, 
		and $R_2$ its stripe.
		Let $\phi$ be the function
		that match $R_2$ to $R_1$.
		We define $\varphi$ as, 
		$\forall x \in [|\pmotif|-1]$,
		$\varphi(x) = \phi(x)$,
		and $\varphi(|\pmotif|) = \phi(|\pmotif|-1)+1$
		Suppose that $\exists x,y \in [|\pmotif|]$, 
		$\pmotif(x)<\pmotif(y)$ 
		and $\ptext(\varphi(x))>\ptext(\varphi(y))$,
		without lost of generality we can suppose that $x<y$,
		Let see all the case possible that can make this 
		hypothesis possible :
		\begin{itemize}		
			\item if $\pmotif(x)<\pmotif(y)$, 
			$\mathcal{S}(\pmotif,x)=\ustep$ 
			and $\mathcal{S}(\pmotif,y)=\ustep$
			(if x and y are increasing in $\pmotif$),
			then $\mathcal{S}(\ptext,\varphi(x))=\ustep$ and $\mathcal{S}(\ptext,\varphi(y))=\ustep$,
			(the match are also increasing).
			As $x<y$ and as $\varphi$ is an increasing function, 
			$\varphi(x) < \varphi(y)$.
			And thus, by \textbf{Lemma \ref{upup}},
			$\ptext(\varphi(x)) < \ptext(\varphi(y))$ .
			Which contradict our hypothesis,
			so this case is not possible.

@@

			\item If $y = |\pmotif|$ and $x=|\pmotif|-1$,
			Clearly if $\mathcal{S}(\pmotif,x)=\ustep$,
			then $\pmotif(y)-\pmotif(x)>0$ 
			and because we have a match in the ribbon
			$\ptext(\varphi(y))-\ptext(\varphi(x))>0$.
			Same idea goes if $\mathcal{S}(\pmotif,x)=\dstep$.
			So this case is not possible.

			\item if $y = |\pmotif|$ and $x<|\pmotif|-1$,
			if $\mathcal{S}(\pmotif,x)=\ustep$ ie
			$\pmotif(x+1)-\pmotif(x)>0$. Then
			we must have either
			$\pmotif(x)<\pmotif(x+1)<\pmotif(|\pmotif|)$
			or $\pmotif(x)<\pmotif(|\pmotif|)<\pmotif(x+1)$ 
			to avoid $231$ or $213$ in the motif.
			Because we have a match in the stripe
			and what we prove above,
			$\ptext(\varphi(x+1)-\ptext(\varphi(x))>0$.
			But by hypothesis $\ptext(\varphi(|\ptext|))<\ptext(\varphi(x))$,
			so we have either 		
			$\pmotif(|\pmotif|)<\pmotif(x)<\pmotif(x+1)$
			or $\pmotif(|\pmotif|)<\pmotif(x)<\pmotif(x+1)$ 
			which both lead to a $231$ or a $213$ motif
			in the text which is not possible.
			The same idea goes if $\mathcal{S}(\pmotif,x)=\dstep$.

@@

			For the algorithm we introduce a decomposition,
			we decompose the motif into
			adjacent element of same step in the stripe 
			(called such sequence segment),
			add the last element to the right most segment,
			and label them form right to left
			with increasing number.
			Let's call this decomposition $S$.
			For each segment, we call important 
			the leftmost element denoted by $IMP$.
			For example, given the motif $(12,11,10,1,2,3,4,9,8,5,6,7)$,
			$S(1)=(5,6,7)$, $IMP(1)=5$, 
			$S(2)=(9,8)$, $IMP(2)=9$, 
			$S(3)=(1,2,3,4)$, $IMP(3)=1$, 
			$S(4)=(12,11,10)$ and $IMP(4)=12$.\\
